---
title: "Gov 50 Final Project"
author: "Julia High"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

INTRODUCTION: 
- Introduce research question
- Introduce hypothesis
- Describe why it is interesting

DATA: 
- Describe the data source
- Describe how the key dependent and independent variables are measured
- Produces a plot that summarizes the dependent variable
- note if research design is cross-sectional or another type

RESULTS: 
- Scatterplot, barplot, or boxplot of the main relationship of interest
- output for the main regression of interest
- interpret the main coefficient of interest in regression
- comment on the statistical significance of the estimated coefficient and whether or not the coefficient represents a causal effect

CONCLUSION: 
- One paragraph that summarizes results
- assesses the extent to which you find support for your hypothesis
- describes limitations of your analysis 
- describes threats to inference
- states how your analysis could be improved


Research question: Do Democrat and Republican views on 5 major social issues differ by racial group?
Hypothesis: Democrat and Republican views on 5 major social issues will differ by racial group.


```{r}
library(tidyverse)
library(dplyr)
```

```{r}
voting_trends_2020 <- read_csv("anes_timeseries_2020_csv_20220210.csv")
voting_trends_2020
```

Race: V201549x
Abortion: V201336
LGBTQ+ Rights: V201416
Immigration: V201417
Environment: V201262
Economy: V201324
Party: V201018 (filter to include democrat and republican)
**maybe I could just do narm = true instead of filtering for the issues. so just filter for race and party!

## I did %in% below to make abortion be 1, 2, 3 or 4
##DO I need to make otehr values lik e that?

```{r}
voting_trends_2020 <- voting_trends_2020 |>
  select(V201549x, V201336, V201416, V201417, V201262, V201324, V201018) |>
  filter(
    V201549x >= 1,
    V201018 == 1 | V201018 == 2,
    V201336 %in% c(1,2,3,4),
    V201416 >= 1,
    V201417 >= 1,
    V201262 %in% c(1, 2, 3, 4, 5, 6, 7),
    V201324 >= 1
  )
voting_trends_2020
```
```{r}
voting_trends <- voting_trends_2020 |>
  rename(
    race_num = V201549x,
    abortion_views = V201336,
    lgbtq_views = V201416,
    immigration_views = V201417,
    environment_views = V201262,
    economy_views = V201324,
    party = V201018
  ) |>
  mutate(
    race = case_when(
      race_num == 1 ~ "White, non-Hispanic",
      race_num == 2 ~ "Black, non-Hispanic",
      race_num == 3 ~ "Hispanic",
      race_num == 4 ~ "Asian or Native Hawaiian/other Pacific Islander, non-Hispanic",
      race_num == 5 ~ "Native American/Alaska Native or other race, non-Hispanic",
      race_num == 6 ~ "Multiple races, non-Hispanic"
    ),
    party = if_else(party == 1, "Democrat", "Republican")
  )
voting_trends
```

IF IT'S DEPENDENT ALONE:

using abortoin_views as the variable
i just need to show the variable and the count
maybe histogram or barplot

```{r}
voting_trends_dependent_abortion <- voting_trends |>
  ggplot(mapping = aes(x = abortion_views)) +
           geom_bar() +
  labs(
    x = "Views on Abortion on a Scale of 1-4",
    y = "Number of Respondents"
  )
voting_trends_dependent_abortion
```
**NEED TO MAKE SURE I DEFINE LABELS AND SCALES

```{r}
voting_trends_dependent_lgbtq <- voting_trends |>
  ggplot(mapping = aes(x = lgbtq_views)) +
           geom_bar() +
  labs(
    x = "Views on Gay Marriage on a Scale of 1-3",
    y = "Number of Respondents"
  )
voting_trends_dependent_lgbtq
```
```{r}
voting_trends_dependent_immigration <- voting_trends |>
  ggplot(mapping = aes(x = immigration_views)) +
           geom_bar() +
  labs(
    x = "Views on Immigration on a Scale of 1-4",
    y = "Number of Respondents"
  )
voting_trends_dependent_immigration
```
```{r}
voting_trends_dependent_environment <- voting_trends |>
  ggplot(mapping = aes(x = environment_views)) +
           geom_bar() +
  labs(
    x = "Views on the Environment on a Scale of 1-7",
    y = "Number of Respondents"
  ) 
voting_trends_dependent_environment
```

**CONTEXTUALIZE survey questions in introduction and write ups

```{r}
voting_trends_dependent_economy <- voting_trends |>
  ggplot(mapping = aes(x = economy_views)) +
           geom_bar() +
  labs(
    x = "Views on the Economy on a Scale of 1-5",
    y = "Number of Respondents"
  )
voting_trends_dependent_economy
```

**is the count here how many people answered with each number? 
or do I summarize them just not group


```{r}
voting_trends_dependent_tibble <- voting_trends |>
  summarize(
    avg_abortion = mean(abortion_views),
    avg_lgbtq = mean(lgbtq_views),
    avg_immigration = mean(immigration_views),
    avg_environment = mean(environment_views),
    avg_economy = mean(economy_views)
  )
knitr::kable(voting_trends_dependent_tibble,
             col.names = c("Average Views on Abortion", "Average Views on Gay Marriage", "Average Views on Immigration", "Average Views on the Environment", "Average Views on the Economy"))
```

**To do this, I would have to make a different plot for each issue since 1-5 means different things



```{r}
voting_tibble <- voting_trends |>
  group_by(party, race) |>
  summarize(
    avg_abortion = mean(abortion_views),
    avg_lgbtq = mean(lgbtq_views),
    avg_immigration = mean(immigration_views),
    avg_environment = mean(environment_views),
    avg_economy = mean(economy_views)
  )
knitr::kable(voting_tibble,
             col.names = c("Party Affiliation", "Race", "Views on Abortion", "Views on Gay Marriage", "Views on Immigration", "Views on the Environment", "Views on the Economy"))
```


```{r}
abortion_plot <- voting_tibble |>
  ggplot(mapping = aes(x = party, y = avg_abortion, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Party Affiliation",
       y = "Average Views on Abortion")
abortion_plot
```

```{r}
lgbtq_plot <- voting_tibble |>
  ggplot(mapping = aes(x = party, y = avg_lgbtq, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Party Affiliation",
       y = "Average Views on Gay Marriage")
lgbtq_plot
```

**MAYBE ADD IN A ROW FOR THE AVERAGE DEM AND AVERAGE REP FOR EACH AS A COMPARISON

```{r}
immigration_plot <- voting_tibble |>
  ggplot(mapping = aes(x = party, y = avg_immigration, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Party Affiliation",
       y = "Average Views on Immigration")
immigration_plot
```

```{r}
environment_plot <- voting_tibble |>
  ggplot(mapping = aes(x = party, y = avg_environment, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Party Affiliation",
       y = "Average Views on the Environment")
environment_plot
```


```{r}
economy_plot <- voting_tibble |>
  ggplot(mapping = aes(x = party, y = avg_economy, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Party Affiliation",
       y = "Average Views on the Economy")
economy_plot
```

```{r}
abortion_regression <- lm(avg_abortion ~ race + party, data = voting_tibble)
summary(abortion_regression)
```
*8Note: These values under RaceBlack, non-Hispanic etc have nothing to do with statistical significance. They mean something else. They are the slope.
slope goes down as compared to what?
since i'm just looking at abortion, the intercept could still be valuable. no results will be valuable if I do an analysis of all the key issues because 1-5 means different things.



```{r}
lgbtq_regression <- lm(avg_lgbtq ~ race + party, data = voting_tibble)
summary(lgbtq_regression)
```

modelsummary::modelsummary(immigration_regression,
statistic = c("s.e. = {std.error}",
"p = {p.value}"),
coef_map = var_labels,
gof_map = c("nobs", "r.squared", "adj.r.squared"))

```{r}
immigration_regression <- lm(avg_immigration ~ race + party, data = voting_tibble)
summary(immigration_regression)
```

```{r}
environment_regression <- lm(avg_environment ~ race + party, data = voting_tibble)
summary(environment_regression)
```

```{r}
economy_regression <- lm(avg_economy ~ race + party, data = voting_tibble)
summary(economy_regression)
```


















