---
title: "Gov 50 Final Project"
author: "Julia High"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---
### INTRODUCTION: 

My research question is: “Do Democrat and Republican views on 5 major social issues differ by racial group?” My hypothesis is that yes, Democrat and Republican views on 5 major social issues will differ by racial group. For this project, the five key social issues I am investigating are abortion, LGBTQ+ rights, immigration, the environment, and the economy. The racial groups I am investigating are “White, non-Hispanic,” “Black, non-Hispanic,” “Hispanic,” “Asian or Native Hawaiian/other Pacific Islander, non-Hispanic,” and “Multiple races, non-Hispanic.” Additionally, I will use the two major parties, Democrats and Republicans, for my analysis. A lot of public opinion research collects respondents’ race, and responses are often analyzed through that lens. Given the history of racial polarization in the United States, analyzing different racial groups’ opinions can reveal interesting contrasts. However, I have not observed many analyses on racial differences within party groups. This is nonetheless significant, given that politicians spend considerable time targeting key racial demographics in elections. This analysis is particularly relevant given the upcoming presidential election in 2024. Additionally, the United States population is rapidly diversifying, so politicians will have no choice but to identify the needs and values of key racial demographic groups. Party players will no longer be able to take advantage of the fact that certain racial groups have historically voted for their party. I hope that my analysis will reveal nuances within party lines and across racial groups that may inspire more comprehensive research in this area. 


### Data - Organizing, Mutating, and Labeling 

My data comes from the American National Election Studies 2020 Time Series study. This source is a public opinion survey of over 8,000 Americans in 2020 about a number of different topics. Party affiliation and race are self-reported in the survey.

I am measuring respondents’ views on my five key social issues based on the following survey questions.

Abortion
Question: There has been some discussion about abortion during recent years. Which one of the opinions on this page best agrees with your view? 
Response options: 1 (By law, abortion should never be permitted) through 4 (By law, a woman should always be able to obtain an abortion as a matter of personal choice).

LGBTQ+ Rights
Question: Which comes closest to your view [on gay marriage]? You can just tell me the number of your choice.
Response options: 1 (Gay and lesbian couples should be allowed to legally marry) through 3 (There should be no legal recognition of gay or lesbian couples’ relationship).

Immigration
Question: Which comes closest to your view about what government policy should be toward unauthorized immigrants now living in the United States?
Response options: 1 (Make all unauthorized immigrants felons and send them back to their home country) through 4 (Allow unauthorized immigrants to remain in US & eventually qualify for citizenship without penalties).

Environment
Question: Where would you place yourself on this scale, or haven’t you thought much about this?
Response options: 1 (Tougher regulations on business needed to protect environment) through 7 (Regulations to protect environment already too much a burden on business).

Economy
Question: What do you think about the state of the economy these days in the United States? Would you say the state of the economy is very good,good, neither good nor bad, bad, or very bad?
Response options: 1(Very good) through 5 (Very bad).

While these categories may not entirely reflect of respondents’ views on a given issue, they offer a solid foundation to reveal potential differences by race and party.

The research design for this study is cross-sectional.

My dependent variable summaries are below, with barplots with the numbers of participants that responded to the survey in a given way for each social issue. 

```{r}
library(tidyverse)
library(dplyr)
```

```{r}
voting_trends_2020 <- read_csv("anes_timeseries_2020_csv_20220210.csv")
```

```{r}
voting_trends_2020 <- voting_trends_2020 |>
  select(V201549x, V201336, V201416, V201417, V201262, V201324, V201018, V201510, V201600) |>
  filter(
    V201549x >= 1,
    V201018 == 1 | V201018 == 2,
    V201336 %in% c(1,2,3,4),
    V201416 >= 1,
    V201417 >= 1,
    V201262 %in% c(1, 2, 3, 4, 5, 6, 7),
    V201324 >= 1,
    V201510 %in% c(1, 2, 3, 4, 5, 6, 7, 8),
    V201600 >= 1
  )
```

```{r}
voting_trends <- voting_trends_2020 |>
  rename(
    race_num = V201549x,
    abortion_views = V201336,
    lgbtq_views = V201416,
    immigration_views = V201417,
    environment_views = V201262,
    economy_views = V201324,
    party = V201018,
    education_level = V201510,
    gender = V201600
  ) |>
  mutate(
    race = case_when(
      race_num == 1 ~ "White, non-Hispanic",
      race_num == 2 ~ "Black, non-Hispanic",
      race_num == 3 ~ "Hispanic",
      race_num == 4 ~ "Asian or Native Hawaiian/other Pacific Islander, non-Hispanic",
      race_num == 5 ~ "Native American/Alaska Native or other race, non-Hispanic",
      race_num == 6 ~ "Multiple races, non-Hispanic"
    ),
    party = if_else(party == 1, "Democrat", "Republican"),
    gender = if_else(gender == 1, "Male", "Female"),
    education_level = case_when(
      education_level == 1 ~ "Less than high school",
      education_level == 2 ~ "High school graduate",
      education_level == 3 ~ "Some college but no degree",
      education_level == 4 ~ "Associate degree in college - vocational",
      education_level == 5 ~ "Associate degree in college - academic",
      education_level == 6 ~ "Bachelor's degree",
      education_level == 7 ~ "Master's degree",
      education_level == 8 ~ "Professional school degree/doctoral degree"
    )
  )
```
### Data - Outcomes of Interest

This Outcome of Interest 1 plot visualizes the counts of respondents who answered 1 - 4 when asked "There has been some discussion about abortion during recent years. Which one of the opinions on this page best agrees with your view?" This plot demonstrates that, regardless of party or race, most respondents chose 4 indicating that they do support a woman's right to bodily autonomy.

```{r}
voting_trends_dependent_abortion <- voting_trends |>
  ggplot(mapping = aes(x = abortion_views)) +
           geom_bar() +
  labs(
    x = "Abortion Views Scale of 1-4",
    y = "Number of Respondents",
    title = "Outcome of Interest 1: All Respondents' Views on Abortion"
  )
voting_trends_dependent_abortion
```

This Outcome of Interest 2 plot visualizes the counts of respondents who answered 1 - 3 when asked "Which comes closest to your view [on gay marriage]?" This plot demonstrates that, regardless of party or race, most respondents chose 1 which suggests that they believe gay and lesbian couples should indeed be allowed to marry.

```{r}
voting_trends_dependent_lgbtq <- voting_trends |>
  ggplot(mapping = aes(x = lgbtq_views)) +
           geom_bar() +
  labs(
    x = "Gay Marriage Views Scale of 1-3",
    y = "Number of Respondents",
    title = "Outcome of Interest 2: All Respondents' Views on Gay Marriage"
  )
voting_trends_dependent_lgbtq
```
This Outcome of Interest 3 plot shows the number of respondents who answered 1 - 4 when asked "Which comes closest to your view about what government policy should be toward unauthorized immigrants now living in the United States?" This plot shows that, without our key independent variables of race and party, most respondents chose 3 which indicates that they believe the US should "Allow unauthorized immigrants to remain in US & eventually
qualify for citizenship but only if they meet requirements."

```{r}
voting_trends_dependent_immigration <- voting_trends |>
  ggplot(mapping = aes(x = immigration_views)) +
           geom_bar() +
  labs(
    x = "Immigration Views Scale of 1-4",
    y = "Number of Respondents",
    title = "Outcome of Interest 3: All Respondents' Views on Immigration"
  )
voting_trends_dependent_immigration
```
This Outcome of Interest 4 plot shows the number of respondents who answered 1 - 7 when asked "Where would you place yourself on this scale, or haven’t you thought much about this?" when asked about the tradeoff between supporting the environment and supporting businesses. This plot shows that, without our key independent variables of race and party, most respondents chose 1 which shows they believe that "Tougher regulations on business [are] needed to protect environment."

```{r}
voting_trends_dependent_environment <- voting_trends |>
  ggplot(mapping = aes(x = environment_views)) +
           geom_bar() +
  labs(
    x = "Environment Views Scale of 1-7",
    y = "Number of Respondents",
    title = "Outcome of Interest 4: All Respondents' Views on the Environment"
  ) 
voting_trends_dependent_environment
```
Finally, this Outcome of Interest 5 plot depicts the number of respondents who answered 1 - 5 when asked "What do you think about the state of the economy these days in the United States? Would you say the state of the economy is very good,good, neither good nor bad, bad, or very bad?" This plot shows that answers ranged widely, but most respondents chose 4 regardless of race or party, which shows that they believe the current state of economy is "bad." Note that this survey was taken in 2020 during a time of significant social and economic turmoil in the United States (and the world) due to the COVID-19 pandemic and racial tensions.

```{r}
voting_trends_dependent_economy <- voting_trends |>
  ggplot(mapping = aes(x = economy_views)) +
           geom_bar() +
  labs(
    x = "Economy Views Scale of 1-5",
    y = "Number of Respondents",
    title = "Outcome of Interest 5: All Respondents' Views on the Economy"
  )
voting_trends_dependent_economy
```

```{r}
library(patchwork)

(voting_trends_dependent_abortion)+(voting_trends_dependent_economy)+(voting_trends_dependent_environment)+(voting_trends_dependent_immigration)+(voting_trends_dependent_lgbtq)
```

### Results - Key Plots 

My results include a table and bar plots of respondents' average views on abortion by party and racial group, as well as regressions of of each social issue with race, party, and the confounders of education level and gender.

```{r}
voting_tibble <- voting_trends |>
  group_by(party, race) |>
  summarize(
    avg_abortion = mean(abortion_views),
    avg_lgbtq = mean(lgbtq_views),
    avg_immigration = mean(immigration_views),
    avg_environment = mean(environment_views),
    avg_economy = mean(economy_views)
  )
knitr::kable(voting_tibble,
             col.names = c("Party Affiliation", "Racial Group", "Average Views on Abortion", "Average Views on Gay Marriage", "Average Views on Immigration", "Average Views on the Environment", "Average Views on the Economy"))
```

Analysis Plot 1 shows respondents' average views on abortion for each party by racial group and party. COME BACK AND EXPLAIN EACH OF THESE. 

```{r}
abortion_plot <- voting_tibble |>
  ggplot(mapping = aes(x = party, y = avg_abortion, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Party Affiliation",
       y = "Average Views on Abortion",
       title = "Analysis Plot 1: Average Views on Abortion by Party Affiliation and Racial Group") 
abortion_plot
```

Analysis Plot 2 shows respondents' average views on gay marriage for each party by racial group and party. 

```{r}
lgbtq_plot <- voting_tibble |>
  ggplot(mapping = aes(x = party, y = avg_lgbtq, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Party Affiliation",
       y = "Average Views on Gay Marriage",
       title = "Analysis Plot 2: Average Views on Gay Marriage by Party Affiliation and Racial Group")
lgbtq_plot
```
Analysis Plot 3 shows respondents' average views on immigration for each party by racial group and party. 

```{r}
immigration_plot <- voting_tibble |>
  ggplot(mapping = aes(x = party, y = avg_immigration, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Party Affiliation",
       y = "Average Views on Immigration",
       title = "Analysis Plot 3: Average Views on Immigration by Party Affiliation and Racial Group")
immigration_plot
```
Analysis Plot 4 shows respondents' average views on the environment for each party by racial group and party. 

```{r}
environment_plot <- voting_tibble |>
  ggplot(mapping = aes(x = party, y = avg_environment, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Party Affiliation",
       y = "Average Views on the Environment",
       title = "Analysis Plot 4: Average Views on the Environment by Party Affiliation and Racial Group")
environment_plot
```
Analysis Plot 5 shows respondents' average views on the economy for each party by racial group and party. 


MAKE SURE TO ROUND NUMBERS
```{r}
economy_plot <- voting_tibble |>
  ggplot(mapping = aes(x = party, y = avg_economy, fill = race)) +
  geom_col(position = "dodge") +
  labs(x = "Party Affiliation",
       y = "Average Views on the Economy",
       title = "Analysis Plot 5: Average Views on the Economy by Party Affiliation and Racial Group")
economy_plot
```

### Results - Key Regressions

```{r}
rep_voting_trends <- voting_trends |>
  filter(
    party == "Republican"
  ) |>
  mutate(
    race3 = case_when(
      race_num == 1 ~ "White, non-Hispanic",
    race_num == 2 ~ "Black, non-Hispanic",
    race_num == 3 ~ "Hispanic",
    race_num == 4 ~ "Asian",
    race_num == 5 ~ "Native American",
    race_num == 6 ~ "Multiple races"
    ))
```

```{r}
dem_voting_trends <- voting_trends |>
  filter(
    party == "Democrat"
  ) |>
  mutate(
    race2 = case_when(
      race_num == 1 ~ "White, non-Hispanic",
    race_num == 2 ~ "Black, non-Hispanic",
    race_num == 3 ~ "Hispanic",
    race_num == 4 ~ "Asian",
    race_num == 5 ~ "Native American",
    race_num == 6 ~ "Multiple races"
    ))
```

For my regressions, I split my voting_trends data into views on each social issue among Democrats and Republicans, respectively. I then created two separate regressions for respondents' views on abortion, LGBTQ+ rights, immigration, the environment, and the economy for each party. In the regression, I analyzed respondents' views as the dependent variable, race as the independent variable, while controlling for eductaional level and gender. My analysis of each of the regressions is below.

This first regression[INSERT TITLE] demonstrates Democrats' average views on abortion in the context of repsondents' race, while controlling for educational level and gender. The second regression shows the same, but for Republican views on abortion. As a reminder, a response of 1 means that abortion should never be permitted, and a response of 4 means that a woman should always be able to obtain an abortion. Thus, responses closer to 1 are the more conservative responses while responses closer to 4 are the more liberal ones.

These two regressions show the following outcomes: 

The first regression [INSERT TITLE] shows that being a Black Democrat decreases your opinion on abortion [AS COMPARED TO OTHER DEMOCRATS?] by -0.093479 while controlling for educational level and gender, while the second shows that being a Black Republican decreases your opinion on abortion by -0.961011 [AS COMPARED TO OTHER REPUBLICANS?] while controlling for those factors. The result for Black Republicans was statistically significant with a threshold of 0.01. This suggests that, when it comes to the abortion issue, identifying as Black may impact Republicans' views (making them more conservative) more than Democrats (instead of making views more liberal). Regardless of race, this result is still consistent how Republicans' average views on abortion were lower for every racial category, including Black, non-Hispanic in Analysis Plot 1. 

**can I say this last sentence? because would I have to say it for the next one? wtf am I comparing to? because technically in the next one republicans are still lower, just decreased less for republicans than decreased for democrats but democrat is still probably higher in general than republican

The first regression shows that being a Hispanic Democrat decreases your opinion on abortion by -0.235807 while controlling for the aforementioned confounders, while the second shows that being a Hispanic Republican decreases your opinion on abortion by -0.147738. This indicates that identifying as Hispanic may impact Democrats' views more than Republicans. The result for Hispanic Democrats was statistically significant with a threshold of 0.01.

The first regression shows that being a Native American Democrat may decrease your views on abortion by -0.150027 with controls, and being a Native American Republican may decrease your views on the issue by -0.132529. When it comes to the issue of abortion, being Native American may impact Democrat views more than Republican views, but only by a small amount.

The first regression shows that being a White Democrat may decrease your views on abortion by -0.003517 while controlling for confounders, while being a White Republican may decrease your views by -0.109134. This suggests that being White has a greater impact on Republican views than Democrat views.

HOW DO I INCLUDE ASIAN AMERICAN IN HERE? do the subtraction i think

**make sure I understand causal
???i'm trying to show the impact of race against the average opinion of that party. the average opinion is the regression that I did the first time.

#### Abortion Regression - Democrats
```{r}
abortion_regression_dem <- lm(abortion_views ~ race2 + education_level + gender, data = dem_voting_trends)
summary(abortion_regression_dem)
modelsummary::modelsummary(abortion_regression_dem,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_rename = c(
                             "(Intercept)" = "Intercept",
                             "race2Black, non-Hispanic" = "Black, non-Hispanic",
                             "race2Hispanic" = "Hispanic",
                             "race2Multiple races" = "Multiple races, non-Hispanic",
                             "race2Native American" = "Native American/Alaska Native or other race, non-Hispanic",
                             "race2White, non-Hispanic" = "White, non-Hispanic",
                             "education_levelAssociate degree in college - vocational" = "Associate's degree in college - vocational",
                             "education_levelBachelor's degree" = "Bachelor's degree",
                             "education_levelHigh school graduate" = "High school graduate",
                             "education_levelLess than high school" = "Less than high school",
                             "education_levelMaster's degree" = "Master's degree",
                             "education_levelProfessional school degree/doctoral degree" = "Professional school/doctoral degree",
                             "education_levelSome college but no degree" = "Some college, but no degree",
                             "genderMale" = "Male"
                             )
                           )
```

#### Abortion Regression - Republicans
```{r}
abortion_regression_rep <- lm(abortion_views ~ race3 + education_level + gender, data = rep_voting_trends)
summary(abortion_regression_rep)
modelsummary::modelsummary(abortion_regression_rep,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_rename = c(
                             "(Intercept)" = "Intercept",
                             "race2Black, non-Hispanic" = "Black, non-Hispanic",
                             "race2Hispanic" = "Hispanic",
                             "race2Multiple races" = "Multiple races, non-Hispanic",
                             "race2Native American" = "Native American/Alaska Native or other race, non-Hispanic",
                             "race2White, non-Hispanic" = "White, non-Hispanic",
                             "education_levelAssociate degree in college - vocational" = "Associate's degree in college - vocational",
                             "education_levelBachelor's degree" = "Bachelor's degree",
                             "education_levelHigh school graduate" = "High school graduate",
                             "education_levelLess than high school" = "Less than high school",
                             "education_levelMaster's degree" = "Master's degree",
                             "education_levelProfessional school degree/doctoral degree" = "Professional school/doctoral degree",
                             "education_levelSome college but no degree" = "Some college, but no degree",
                             "genderMale" = "Male"
                             )
                           )
```

The first regression [INSERT TITLE] shows Democrats' average views on gay marriage in the context of the respondents' race, while controlling for educational level and gender. The second regression shows the same, but for Republican views on gay marriage. In this case, a response of 1 means that a respondent supports gay marriage while a response of 3 means that a respondent does not support it. Therefore, as responses decrease and get closer to 1, they are the more liberal view and while responses increase closer to 3 they are more conservative.

These two regressions show the following outcomes: 

The first regression [INSERT TITLE] shows that being a Black Democrat may increase respondents' opinion on gay marriage by 0.23674 while controlling for educational level and gender. The second shows that being a Black Republican may increase your opinion on gay marriage by 0.357982 with the same controls. This suggests that identifying as Black may have a greater impact on Republican respondents' views on gay marriage than Democrat respondents. 

In terms of respondents who identified as Hispanic, the first regression shows that being a Hispanic Democrat increases respondents' views on gay marriage by 0.05924 while controlling for educational level and gender, while being a Hispanic Republican decreases those views by -0.255243. This suggests that being Hispanic has a larger impact on Republicans than Democrats when it comes to views on gay marriage. 

The first regression shows that being a multiple races and a Democrat may decrease respondents' opinions on gay marriage by -0.06999. Being multiple races and a Republican, on the other hand, may decrease opinions on abortion by -0.283640. This indicates that being multiple races may impact Republican views more than Democrat views on this issue, which aligns with what we observed with the previous two racial groups.

The first regression also shows that being a Native American Democrat may decrease respondents' opinions on gay marriage by -0.11899 while controlling for educational level and gender, and the second regression shows that being a Native American Republican may decrease those opinions by -0.304337. This suggests that being Native American may have a greater impact on Republican respondents than Democrats once again when it comes to the issue of gay marriage.

Finally, the first regression shows that being a White Democrat may decrease respondents' views on gay marriage by -0.04897 while controlling for the aforementioned factors and being a White Republican may decrease their views on the same issue by -0.241769. This suggests that being White may have more of an impact on Republican respondents' responses in regards to gay marriage.

#### LGBTQ Regression - Democrats
```{r}
lgbtq_regression_dem <- lm(lgbtq_views ~ race2 + education_level + gender, data = dem_voting_trends)
summary(lgbtq_regression_dem)
modelsummary::modelsummary(lgbtq_regression_dem,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_rename = c(
                             "(Intercept)" = "Intercept",
                             "race2Black, non-Hispanic" = "Black, non-Hispanic",
                             "race2Hispanic" = "Hispanic",
                             "race2Multiple races" = "Multiple races, non-Hispanic",
                             "race2Native American" = "Native American/Alaska Native or other race, non-Hispanic",
                             "race2White, non-Hispanic" = "White, non-Hispanic",
                             "education_levelAssociate degree in college - vocational" = "Associate's degree in college - vocational",
                             "education_levelBachelor's degree" = "Bachelor's degree",
                             "education_levelHigh school graduate" = "High school graduate",
                             "education_levelLess than high school" = "Less than high school",
                             "education_levelMaster's degree" = "Master's degree",
                             "education_levelProfessional school degree/doctoral degree" = "Professional school/doctoral degree",
                             "education_levelSome college but no degree" = "Some college, but no degree",
                             "genderMale" = "Male"
                             )
                           )
```

#### LGBTQ Regression - Republicans
```{r}
lgbtq_regression_rep <- lm(lgbtq_views ~ race3 + education_level + gender, data = rep_voting_trends)
summary(lgbtq_regression_rep)
modelsummary::modelsummary(lgbtq_regression_rep,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_rename = c(
                             "(Intercept)" = "Intercept",
                             "race2Black, non-Hispanic" = "Black, non-Hispanic",
                             "race2Hispanic" = "Hispanic",
                             "race2Multiple races" = "Multiple races, non-Hispanic",
                             "race2Native American" = "Native American/Alaska Native or other race, non-Hispanic",
                             "race2White, non-Hispanic" = "White, non-Hispanic",
                             "education_levelAssociate degree in college - vocational" = "Associate's degree in college - vocational",
                             "education_levelBachelor's degree" = "Bachelor's degree",
                             "education_levelHigh school graduate" = "High school graduate",
                             "education_levelLess than high school" = "Less than high school",
                             "education_levelMaster's degree" = "Master's degree",
                             "education_levelProfessional school degree/doctoral degree" = "Professional school/doctoral degree",
                             "education_levelSome college but no degree" = "Some college, but no degree",
                             "genderMale" = "Male"
                             )
                           )
```

The following regressions depict respondents' average views on immigration in the context of the respondents' race, while controlling for educational level and gender. The first regression represents the average views of Democrat respondents, while the second represents the views of Republican respondents. In this case, a response of 1 means that the respondent believes the US should make unauthorized immigrants felons and send them back to their home country, and a response of 4 means that respondents believe unauthorized immigrants should be allowed to remain in the US and eventually qualify for citizenship without penalties. This means that responses with lower values that are closer to 1 are more conservative, while responses with higher values that are closer to 4 are more liberal.

The two regressions show the following outcomes: 

The first regression [INSERT TITLE] shows that being a Black Democrat may increase respondents' views on immigration by 0.07495 while controlling for educational level and gender, while the second regression shows that being a Black Republican may decrease respondents' views on immigration by -0.109081 while controlling for the same factors. This suggests that being Black has a greater impact on Republican responses than Democrat responses.

The first regression shows that being a Hispanic Democrat may increase respondents' views on immigration by 0.32177 while controlling for confounders. This result was also statistically significant with a threshold of 0.001. The second regression, on the other hand, shows that being a Hispanic Republican may decrease respondents' views on immigration by -0.041310. This indicates that being Hispanic may have a greater impact on Republican views on immigration than Democrats.

[NOTE THAT IM REMOVING MULTIPLE RACES]

The first regression depicts that being a Native American Democrat may increase respondents' views on immigration by 0.46549 while controlling for educational level and gender. This result was statistically significant with a threshold of 0.001. The second regression suggests that being a Native American Republican may decrease respondents' views on immigration by -0.192312. In this case, these coefficients suggest that being a Native American may impact Democrats significantly more than Republicans on the issue of immigration.

The first regression indicates that being a White Democrat may increase respondents' views on immigration by 0.25649. This result was statistically significant under a 0.001 threshold. The second regression suggests that being a White Republican may correspond with a decreased view on immigration of -0.118284. When it comes to the issue of immigration, it seems that identifying as White may correspond with a greater impact on Democrats than Republicans.

#### Immigration Regression - Democrats
```{r}
immigration_regression_dem <- lm(immigration_views ~ race2 + education_level + gender, data = dem_voting_trends)
summary(immigration_regression_dem)
modelsummary::modelsummary(immigration_regression_dem,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_rename = c(
                             "(Intercept)" = "Intercept",
                             "race2Black, non-Hispanic" = "Black, non-Hispanic",
                             "race2Hispanic" = "Hispanic",
                             "race2Multiple races" = "Multiple races, non-Hispanic",
                             "race2Native American" = "Native American/Alaska Native or other race, non-Hispanic",
                             "race2White, non-Hispanic" = "White, non-Hispanic",
                             "education_levelAssociate degree in college - vocational" = "Associate's degree in college - vocational",
                             "education_levelBachelor's degree" = "Bachelor's degree",
                             "education_levelHigh school graduate" = "High school graduate",
                             "education_levelLess than high school" = "Less than high school",
                             "education_levelMaster's degree" = "Master's degree",
                             "education_levelProfessional school degree/doctoral degree" = "Professional school/doctoral degree",
                             "education_levelSome college but no degree" = "Some college, but no degree",
                             "genderMale" = "Male"
                             )
                           )
```

#### Immigration Regression - Republicans
```{r}
immigration_regression_rep <- lm(immigration_views ~ race3 + education_level + gender, data = rep_voting_trends)
summary(immigration_regression_rep)
modelsummary::modelsummary(immigration_regression_rep,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_rename = c(
                             "(Intercept)" = "Intercept",
                             "race2Black, non-Hispanic" = "Black, non-Hispanic",
                             "race2Hispanic" = "Hispanic",
                             "race2Multiple races" = "Multiple races, non-Hispanic",
                             "race2Native American" = "Native American/Alaska Native or other race, non-Hispanic",
                             "race2White, non-Hispanic" = "White, non-Hispanic",
                             "education_levelAssociate degree in college - vocational" = "Associate's degree in college - vocational",
                             "education_levelBachelor's degree" = "Bachelor's degree",
                             "education_levelHigh school graduate" = "High school graduate",
                             "education_levelLess than high school" = "Less than high school",
                             "education_levelMaster's degree" = "Master's degree",
                             "education_levelProfessional school degree/doctoral degree" = "Professional school/doctoral degree",
                             "education_levelSome college but no degree" = "Some college, but no degree",
                             "genderMale" = "Male"
                             )
                           )
```

The first regression demonstrates Democrats' average views on the environment in the context of respondents' race while also controlling for educational level and gender. The second regression shows the same, but for Republican views on the environment. In regards to the issue of environment, a response that is decreasing or is closer to 1 means that tougher business regulations are needed to protect the environment, and responses that increase closer to 7 mean that regulations to the environment are already too much of a burden on businesses. This means that responses closer to 1 are more liberal, while responses closer to 7 are more conservative.

The two regressions on the environment show the following outcomes: 

The first regression [INSERT TITLE] suggests that being a Black Democrat decreases your views on the environment by -0.25855 while controlling for the confounders of educational level and gender. The second regression depicts that being a Black Republican increases your views on the environment by 0.48109. In this case, the coefficients suggest that being Black impacts Republican respondents' views more than Democrats.

The first regression indicates that being a Hispanic Democrat decreases your views on the environment by -0.68899, and this result is statistically significant with a threshold of 0. The second regression suggests that being a Hispanic Republican increases your views on the environment by 0.58712. In this scenario, the coefficients suggest that being Hispanic has a greater impact on Democrat respondents than Republicans.

The first regression depicts that being a Native American Democrat decreases your views on the environment by -0.59861. The second regression suggests that being a Native American Republican increases your views on the environment by 0.60353. This suggests that being Native American may have a slightly greater impact on Republican respondents than on Democrat respondents.

The first regression suggests that being a White Democrat decreases your views on the environment by -0.60597, and this result is statistically significant with a threshold of 0. The second regression indicates that being a White Republican may increase your views on the environment by 0.64241. These coefficients indicate that identifying as White may have a slightly greater impact on Democrat than Republican respondents. 

**round all numbers
***GO BACK THROUGH FOR STATISTICAL SIGNIFICANCE
**CHANGE ALL SHOWS TO SUGGESTS

#### Environment Regression - Democrats
```{r}
environment_regression_dem <- lm(environment_views ~ race2 + education_level + gender, data = dem_voting_trends)
summary(environment_regression_dem)
modelsummary::modelsummary(environment_regression_dem,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_rename = c(
                             "(Intercept)" = "Intercept",
                             "race2Black, non-Hispanic" = "Black, non-Hispanic",
                             "race2Hispanic" = "Hispanic",
                             "race2Multiple races" = "Multiple races, non-Hispanic",
                             "race2Native American" = "Native American/Alaska Native or other race, non-Hispanic",
                             "race2White, non-Hispanic" = "White, non-Hispanic",
                             "education_levelAssociate degree in college - vocational" = "Associate's degree in college - vocational",
                             "education_levelBachelor's degree" = "Bachelor's degree",
                             "education_levelHigh school graduate" = "High school graduate",
                             "education_levelLess than high school" = "Less than high school",
                             "education_levelMaster's degree" = "Master's degree",
                             "education_levelProfessional school degree/doctoral degree" = "Professional school/doctoral degree",
                             "education_levelSome college but no degree" = "Some college, but no degree",
                             "genderMale" = "Male"
                             )
                           )
```

#### Environment Regression - Republicans
```{r}
environment_regression_rep <- lm(environment_views ~ race3 + education_level + gender, data = rep_voting_trends)
summary(environment_regression_rep)
modelsummary::modelsummary(environment_regression_rep,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_rename = c(
                             "(Intercept)" = "Intercept",
                             "race2Black, non-Hispanic" = "Black, non-Hispanic",
                             "race2Hispanic" = "Hispanic",
                             "race2Multiple races" = "Multiple races, non-Hispanic",
                             "race2Native American" = "Native American/Alaska Native or other race, non-Hispanic",
                             "race2White, non-Hispanic" = "White, non-Hispanic",
                             "education_levelAssociate degree in college - vocational" = "Associate's degree in college - vocational",
                             "education_levelBachelor's degree" = "Bachelor's degree",
                             "education_levelHigh school graduate" = "High school graduate",
                             "education_levelLess than high school" = "Less than high school",
                             "education_levelMaster's degree" = "Master's degree",
                             "education_levelProfessional school degree/doctoral degree" = "Professional school/doctoral degree",
                             "education_levelSome college but no degree" = "Some college, but no degree",
                             "genderMale" = "Male"
                             )
                           )
```


These two regressions depict survey respondents' average views on the economy in the context of respondents' race while also controlling for educational level and gender. The first regression includes Democrat respondents' views, while the second includes Republican respondents' views. When it comes to the economy, a response that is closer to 1 means that respondents believe the economy is doing very well, and a response that is closer to 5 means that respondents believe the economy is doing very badly. Given the survey question, there is no typical "liberal" or "conservative" response to this question. This analysis will simply show the difference in views on the question by party and racial group, not necessarily which direction those views trend on the political spectrum.

The two regressions show the following outcomes:

The first regression [INSERT TITLE] suggests that being a Black Democrat increases respondent views on the economy by 0.42789, and this coefficient is statistically significant with a 0.001 threshold. The second regression suggests that being a Black Republican decreases your views on the economy by -0.53291. Comparing these coefficients shows that when it comes to the economy, being a Black impacts Republican respondents more than Democrat respondents.

The first regression shows that being a Hispanic Democrat may increase respondent views on the economy by 0.44569. This coefficient is statistically significant with a 0.001 threshold. The second regression suggests that being a Hispanic Republican may decrease your views on the environment by -0.76620. This coefficient is also statistically significant with a 0.001 threshold. Comparing these coefficients shows that being Hispanic may impact Republican respondents more than Democrat respondents on the economy question.

The first regression illustrates that being a Native American Democrat may increase respondent views on the economy by 0.74606. This coefficient is statistically significant with a 0.001 threshold. The second regression illustrates that being a Native American Republican may decrease respondent views on the economy by -1.03236. This result is statistically significant with a threshold of 0. Together, these coefficients indicate that being Native American has a greater impact on Republican responses to the economy question than Democrat responses.

The first regression shows that being a White Democrat may increase respondent views on the economy by 0.35694. This coefficient is statistically significant with a 0.001 threshold. The second regression indicates that being a White Republican may decrease respondent opinions on the economy by -0.77982, and this coefficient is statistically significant with 0 as a threshold. Comparing these coefficients indicates that being White has a greater impact on Republican respondents than Democrats.

#### Economy Regression - Democrats 
```{r}
economy_regression_dems <- lm(economy_views ~ race2 + education_level + gender, data = dem_voting_trends)
summary(economy_regression_dems)
modelsummary::modelsummary(economy_regression_dems,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_rename = c(
                             "(Intercept)" = "Intercept",
                             "race2Black, non-Hispanic" = "Black, non-Hispanic",
                             "race2Hispanic" = "Hispanic",
                             "race2Multiple races" = "Multiple races, non-Hispanic",
                             "race2Native American" = "Native American/Alaska Native or other race, non-Hispanic",
                             "race2White, non-Hispanic" = "White, non-Hispanic",
                             "education_levelAssociate degree in college - vocational" = "Associate's degree in college - vocational",
                             "education_levelBachelor's degree" = "Bachelor's degree",
                             "education_levelHigh school graduate" = "High school graduate",
                             "education_levelLess than high school" = "Less than high school",
                             "education_levelMaster's degree" = "Master's degree",
                             "education_levelProfessional school degree/doctoral degree" = "Professional school/doctoral degree",
                             "education_levelSome college but no degree" = "Some college, but no degree",
                             "genderMale" = "Male"
                             )
                           )
```

#### Economy Regression - Republicans
```{r}
economy_regression_rep <- lm(economy_views ~ race3 + education_level + gender, data = rep_voting_trends)
summary(economy_regression_rep)
modelsummary::modelsummary(economy_regression_rep,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_rename = c(
                             "(Intercept)" = "Intercept",
                             "race2Black, non-Hispanic" = "Black, non-Hispanic",
                             "race2Hispanic" = "Hispanic",
                             "race2Multiple races" = "Multiple races, non-Hispanic",
                             "race2Native American" = "Native American/Alaska Native or other race, non-Hispanic",
                             "race2White, non-Hispanic" = "White, non-Hispanic",
                             "education_levelAssociate degree in college - vocational" = "Associate's degree in college - vocational",
                             "education_levelBachelor's degree" = "Bachelor's degree",
                             "education_levelHigh school graduate" = "High school graduate",
                             "education_levelLess than high school" = "Less than high school",
                             "education_levelMaster's degree" = "Master's degree",
                             "education_levelProfessional school degree/doctoral degree" = "Professional school/doctoral degree",
                             "education_levelSome college but no degree" = "Some college, but no degree",
                             "genderMale" = "Male"
                             )
                           )
```


### Conclusion
	
	**EXPLAIN THIS AND MY DATASET MORE
	**EXPLAIN THAT DATASET DIDN'T INCLUDE EVERYTHING I WANTED
	add a section called tidying and mutating data
	change colors on dependent graphs
	add lines at the bottom of my sections
	show the main variable regression and multiple regressoins
	add works cited
	learn to bold section titles
	**note that I changed my whole project from the milestones





### Works Cited

American National Election Studies. 2021. ANES 2020 Time Series
Study Full Release [dataset and documentation]. July 19, 2021
version. www.electionstudies.org
